---
title: "Convert Figures to JPEG for JEB Submission"
subtitle: "Pure format conversion — does NOT rerun analysis or change figure content"
author: Shannon N. Snyder
date: "`r Sys.Date()`"
format:
  html:
    self-contained: true
    toc: true
execute:
  echo: true
  message: true
  warning: false
---

# Overview

This document converts all manuscript PNG figures to JPEG format at 800 DPI for submission
to the Journal of Evolutionary Biology.

**What this does:**

- Reads each existing `.png` figure from `01.figures/final_revision_JEB_submission/`
- Writes an identical `.jpg` version to `01.figures/final_revision_JEB_submission_jpeg/`
- Sets DPI metadata to 800 in the output file
- Does **not** rerun any statistical analysis or alter figure content
- Figures 1 and 2 are handled separately (Figure1 is PNG, Figure2 is PDF already)

**Accepted JEB formats:** `.jpeg`, `.jpg`, `.tiff`, `.pdf`, `.eps`

---

# Setup

```{r}
#| label: setup

library(magick)
library(here)

# Output directory
out_dir <- "../01.figures/final_revision_JEB_submission_jpeg"
dir.create(out_dir, showWarnings = FALSE, recursive = TRUE)
cat("Output directory:", out_dir, "\n")
```

---

# Figure 1 (PNG → JPEG)

Figure 1 is a schematic imported as `Figure1@4x.png`. Converted here.

```{r}
#| label: convert-fig1

fig1_in  <- "../01.figures/final_revision_JEB_submission/Figure1@4x.png"
fig1_out <- file.path(out_dir, "Figure1.jpg")

img <- image_read(fig1_in)
img_info <- image_info(img)
cat("Input:  ", fig1_in, "\n")
cat("Pixels: ", img_info$width, "x", img_info$height, "\n")
cat("DPI in: ", img_info$density, "\n")

img_out <- image_convert(img, format = "jpeg")
img_out <- image_write(img_out, path = fig1_out, quality = 95, density = "800")
image_info(image_read(fig1_out))[, c("format","width","height","density")]
cat("Saved →", fig1_out, "\n")
```

---

# Figure 2 (PDF — already accepted format)

Figure 2 is already a `.pdf`, which is accepted by JEB. It is copied as-is.

```{r}
#| label: copy-fig2

fig2_in  <- "../01.figures/final_revision_JEB_submission/Figure2.pdf"
fig2_out <- file.path(out_dir, "Figure2.pdf")
file.copy(fig2_in, fig2_out, overwrite = TRUE)
cat("Copied →", fig2_out, "(PDF — no conversion needed)\n")
```

---

# Figures 3–7 (PNG → JPEG)

```{r}
#| label: convert-figs3-7

main_figs <- list(
  list(in_file = "Figure3_ontogenetic_trajectories_revised.png",
       out_file = "Figure3_ontogenetic_trajectories_revised.jpg"),
  list(in_file = "Figure4_predicted_spines_revised.png",
       out_file = "Figure4_predicted_spines_revised.jpg"),
  list(in_file = "Figure5_growth_rates_revised.png",
       out_file = "Figure5_growth_rates_revised.jpg"),
  list(in_file = "Figure6_size_selective_mortality_revised.png",
       out_file = "Figure6_size_selective_mortality_revised.jpg"),
  list(in_file = "Figure7_survival_GAM_revised.png",
       out_file = "Figure7_survival_GAM_revised.jpg")
)

for (fig in main_figs) {
  in_path  <- paste0("../01.figures/final_revision_JEB_submission/", fig$in_file)
  out_path <- file.path(out_dir, fig$out_file)

  img     <- image_read(in_path)
  inf     <- image_info(img)
  img_out <- image_convert(img, format = "jpeg")
  image_write(img_out, path = out_path, quality = 95, density = "800")

  inf_out <- image_info(image_read(out_path))
  cat(sprintf("%-55s  %dx%d px  →  %s\n",
              fig$in_file, inf$width, inf$height, out_path))
}
```

---

# Supplemental Figures S1–S7 (PNG → JPEG)

```{r}
#| label: convert-supp-figs

supp_figs <- list(
  list(in_file = "FigureS1_age_bodylength_revised.png",
       out_file = "FigureS1_age_bodylength_revised.jpg"),
  list(in_file = "FigureS2_headspine_diagnostics_revised.png",
       out_file = "FigureS2_headspine_diagnostics_revised.jpg"),
  list(in_file = "FigureS3_tailspine_diagnostics_revised.png",
       out_file = "FigureS3_tailspine_diagnostics_revised.jpg"),
  list(in_file = "FigureS4_linear_predictions_revised.png",
       out_file = "FigureS4_linear_predictions_revised.jpg"),
  list(in_file = "FigureS5_raw_spine_values_revised.png",
       out_file = "FigureS5_raw_spine_values_revised.jpg"),
  list(in_file = "FigureS6_predicted_spines_with_raw_revised.png",
       out_file = "FigureS6_predicted_spines_with_raw_revised.jpg"),
  list(in_file = "FigureS7_bodylength_diagnostics_revised.png",
       out_file = "FigureS7_bodylength_diagnostics_revised.jpg")
)

for (fig in supp_figs) {
  in_path  <- paste0("../01.figures/final_revision_JEB_submission/", fig$in_file)
  out_path <- file.path(out_dir, fig$out_file)

  img     <- image_read(in_path)
  inf     <- image_info(img)
  img_out <- image_convert(img, format = "jpeg")
  image_write(img_out, path = out_path, quality = 95, density = "800")

  cat(sprintf("%-60s  %dx%d px  →  %s\n",
              fig$in_file, inf$width, inf$height, out_path))
}
```

---

# Verification

Confirm all expected output files exist and are valid JPEGs.

```{r}
#| label: verify

expected <- c(
  "Figure1.jpg",
  "Figure2.pdf",
  "Figure3_ontogenetic_trajectories_revised.jpg",
  "Figure4_predicted_spines_revised.jpg",
  "Figure5_growth_rates_revised.jpg",
  "Figure6_size_selective_mortality_revised.jpg",
  "Figure7_survival_GAM_revised.jpg",
  "FigureS1_age_bodylength_revised.jpg",
  "FigureS2_headspine_diagnostics_revised.jpg",
  "FigureS3_tailspine_diagnostics_revised.jpg",
  "FigureS4_linear_predictions_revised.jpg",
  "FigureS5_raw_spine_values_revised.jpg",
  "FigureS6_predicted_spines_with_raw_revised.jpg",
  "FigureS7_bodylength_diagnostics_revised.jpg"
)

results <- data.frame(
  file   = expected,
  exists = file.exists(file.path(out_dir, expected)),
  size_kb = sapply(file.path(out_dir, expected), function(f) {
    if (file.exists(f)) round(file.size(f) / 1024, 1) else NA
  })
)

print(results, row.names = FALSE)

n_ok  <- sum(results$exists)
n_exp <- nrow(results)
cat(sprintf("\n%d / %d files present in %s\n", n_ok, n_exp, out_dir))

if (n_ok == n_exp) {
  cat("All figures converted successfully. Ready for JEB submission.\n")
} else {
  missing <- results$file[!results$exists]
  cat("MISSING files:\n")
  cat(paste0("  - ", missing, "\n"))
}
```
